---
title: "HW2 Prince George's County Crime"
---

In this assignment, we will be working with a dataset reporting the number of crimes by month in Prince George's County, Maryland, for every month in 2024. 

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv. You will create a dataframe called `pg_crime` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
pg_crime <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/pg_crime_totals_2024.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, are any of the categories more important than others? Why or why not? Do any of them make the job of summarizing the data more difficult? Why or why not?

The data shows 18-types of crimes that have happened in Prince George's County throughout the year in 2024.Some examples of the categories that are listed in the data are "ASSULT WITH WEAPON," "HOMICIDE," and "SEX OFFENESE." Some of these types of crimes will make it hard to report without speaking with the records custodian. For example, we know that homicides involve the murder of a person, but what does that say about assault with weapon or assault with gun. Is it possible for the person to have been assaulted and died in the hospital later. Another problem will be looking at the B & E cases. If B & Es are tracked whether commercial or residential property was broken into, what is the B & E other section? More questions need to be asked by the records keeper.

### Task 4: Calculating Averages

Add up the total number of incidents per month for each category using `mutate()`, calling the new column `total_incidents`. Then calculate the average number of incidents per month for each category of crime using the total number of months, calling the new column `average_incidents`. Assign the result to a new dataframe called `pg_crime_averages`.

```{r}
pg_crime_averages <- pg_crime |> mutate(total_incidents = (Jan + Feb + Mar + Apr + May + Jun + Jul + Aug + Sep + Oct + Nov + Dec)) |> mutate(average_incidents = (total_incidents / 12))
```

### Task 5: Describe the Averages

Examine the new dataframe, `pg_crime_averages`, by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the average number of incidents per month for each category. Which categories have the highest average number of incidents per month? Which have the lowest? What does this tell you about the pattern of crime in Prince George's County? Are there categories that you have questions about? Why or why not?

The three crimes with the largest averages are "ACCIDENT," "THEFT FROM AUTO," and "AUTO STOLEN."
"HOMOCIDE" AND "B&E OTHER" have the lowest averages. At face value, this tells me that the homicide rate in Prince George's County is low, but my journalism brain is telling me that there possibly could be other counts of homicide that may have been added in other categories as I discussed up above.


### Task 6: Testing a Theory

One theory about crime is that it is more likely to occur in the summer months. Test this theory by creating three new columns using `mutate()` that calculate the percentage change compared to the average for each category of crime for the months of June, July, and August. Call the new columns `june_change`, `july_change`, and `august_change`. Assign the result to a new dataframe called `summer_crime_averages`.

```{r}
summer_crime_averages <- pg_crime_averages |> mutate(june_change = (Jun - average_incidents / average_incidents * 100)) |> mutate(july_change = (Jul - average_incidents / average_incidents * 100)) |> mutate(august_change = (Aug - average_incidents / average_incidents * 100))
```

### Task 7: Describe the Summer Crime Averages

Look at the `summer_crime_averages` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change in the average number of incidents per month for each category of crime for the months of June, July, and August. Do you see any patterns? What does this tell you about the theory that crime is more likely to occur in the summer months?

The data shows that only some crimes happen more frequently in the summer months.

### Task 8: What's the Story?

Write up a lede and nutgraf of a story focused on the crime data in Prince George's County. It could be from any of the dataframes you've created in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

Prince Gregory's County sees massive jumps in thefts and accidents in 2024, according to the county's crime data.

Theft from a car increased 250% during the months of June, July and August while accidents nearly increased 400% during these months. Auto thefts also nearly passed 300% during the summer.


### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
