---
title: "HW8_MandMs"
---

In this assignment, you'll be picking up where you left off with the M&Ms activity. You'll load the data and perform some basic statistical analysis. Follow the steps below to complete the assignment.

### Task 1: Load the tidyverse

```{r}
library(tidyverse)
```



### Task 2: Load the M&Ms dataset from https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv

```{r}
MM_data <- read_csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/mandms.csv")
```

### Task 3: Sample 5 rows of data from the dataset

```{r}
set.seed(40)
sample_5 <- MM_data |> sample_n(5)
#mean(sample_50$unaffiliated)
```


### Task 4: Calculate the mean age and standard deviation of the sample, using either the red or black number of M&Ms

```{r}
mean(sample_5$red)
sd(sample_5$red)
```


### Task 5: Calculate the mean age and standard deviation of the population, using the same measure as in Task 4

```{r}
mean(MM_data$total)
sd(MM_data$total)
```


### Task 6: Compare the sample and population statistics

Compare the two sets of means and standard deviations. Are they similar?

The two means are very different. The population mean is much larger than the sample size's mean. The sample and puopulation have similar standard deviations with only 7 points apart. 


### Task 7: Describe the results

Write a few sentences about what you found in the data. What does this tell you about the M&Ms dataset and about the act of random sampling?

All of the total are over 100, which let's me know that the scoops were not controlled, giving us a random samples.

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
